\documentclass{article}


\usepackage[margin=1in]{geometry}
\begin{document}

\title{Intro to Finite Elements Method}
\date{}
\author{Martin Jones}
\maketitle
\thispagestyle{empty}
It is possible to do calculations on an arbitrary grid of finite elements given just one master element. The idea is useful because performing calculations on arbitrary elements in curvilinear coordinates proves awkward, and the character of these calculations such as limits of integration, etc..., changes from element to element.
We call the Master element, $\hat{\Omega}$, and calculations for the mesh we are interested in are performed for each individual element on it, then `transformed' back to the original mesh coordinates, with their respective material properties, or solution, in place.

\begin{equation} 
\frac{\partial \phi}{\partial t} = \vec{\nabla} \cdot (D(\vec{x})\vec{\nabla} \phi)
\end{equation} 
Where $\phi$ and $D$ are concentration and diffusivity, repectively, scalar functions that depend on space and time. And we have $\phi(\vec{x},0)=\phi_{0}(\vec{x})$ and $\vec{\nabla}\phi \cdot \vec{\mathbf{n}}=0$ on the surface for our boundary conditions.

Now multiply both sides by an arbitrary test function $V(x)$ and integrate over a volume $\Omega$. Also, integrating by parts on the RHS, the equation evolves to the following form,  (making use first of the following theorem):

\begin{displaymath}
\int_{\Omega}V\vec{\nabla}\cdot(D\vec{\nabla}\phi)d\Omega=\int_{S}V(D\vec{\nabla}\phi\cdot \vec{\mathbf{n}})dS-\int_{\Omega}D(\vec{\nabla}V\cdot \vec{\nabla} \phi)d\Omega
\end{displaymath}

 \begin{equation} 
\int_{\Omega} \frac{\partial \phi}{\partial t} V(x)d\Omega = -\int_{\Omega}D(\vec{x})\vec{\nabla}V \cdot\vec{\nabla} \phi d\Omega
\end{equation}
Now we will discretize and write with basis functions in N dimensional space (using Einstein notation) where N is on the order of the number of elements in the mesh,

\begin{equation}
\phi (\vec{x},t_{k})=\phi_{i}^{k}\psi_{i}
\end{equation}and
\begin{equation}
V(\vec{x})=V_{j}\psi_{j}
\end{equation}

Using the finite difference formula for the time derivative and plugging in we get,

 \begin{equation} 
\int_{\Omega} \left(\frac{\phi_{i}^{k+1}-\phi_{i}^{k}}{\Delta t}\right)\cdot \psi_{i} (V_{j}\cdot \psi_{j })d\Omega = -\int_{\Omega}D(\vec{x})\vec{\nabla}(V_{j}\psi_{j}) \cdot\vec{\nabla} (\phi_{i}^{k+1}\psi_{i}) d\Omega
\end{equation}
Now we may write N equations with N unknowns due to the arbitrariness of V (1 on the $i^{th}$ node and zero every where else). One gets,

\begin{equation}
\left(\frac{1}{\Delta t}\vec{M}+\vec{K}\right)\vec{\phi}^{k+1}=\frac{1}{\Delta t}\vec{M}\vec{\phi}^{k}
\end{equation}
Where M is given by,
\begin{equation}
M_{i,j}=\int_{\Omega}\psi_{i}\psi_{j}d\Omega
\end{equation}
and K is given by,
\begin{equation}
K_{i,j}=\int_{\Omega}D(\vec{x})\vec{\nabla}\psi_{i}\cdot \vec{\nabla}\psi_{j}d\Omega
\end{equation}

These equations describe the physics on the local grid. We want to describe them on the global grid, or our final mesh. Remembering the jacobian relating the differentials between the two coordinate systems: $(x,y)$, and $(\xi,\eta)$,
\begin{equation}
dx=\frac{\partial x}{\partial \xi}d \xi + \frac{\partial x}{\partial \eta}d \eta
\end{equation}
and,
\begin{equation}
dy=\frac{\partial y}{\partial \xi}d \xi + \frac{\partial y}{\partial \eta}d \eta
\end{equation}
We can obtain our final equations through these equations ultimately. They are.
\begin{equation}
k_{ij}^e=\int_{\Omega_e}\left[ k\left( \frac{\partial \psi_i^e}{\partial x}\frac{\partial \psi_j^e}{\partial x} + b\psi_i^e \psi_j^e  \right) \right]dxdy
\end{equation}
\begin{equation}
f_i^e=\int_{\Omega_e}f\psi_i^edxdy
\end{equation}
\begin{equation}
P_{ij}^e=\int_{\delta \Omega^e}p\psi_i^e\psi_j^e ds
\end{equation}
\begin{equation}
\gamma_i^e = \int_{\delta \Omega^e} \gamma \psi_i^e ds
\end{equation}

\end{document}













